---
title: "Assignment |Reproducible Research Week 2"
author: "Juan M Sanchez"
date: "6/22/2020"
output: html_document
---

### QUESTION 1| Code for reading in the dataset and/or processing the data. 

```{r echo=FALSE}
setwd("/Users/juan7sanchez/Desktop/BI/data")
```
I read and called data frame 'act'. 

```{r}
act<- read.csv(file='activity.csv', header = TRUE)
```

I checked data frame and ajust variable class: 

```{r}
str(act)
act$steps <- as.numeric(as.character(act$steps))
act$interval <- as.numeric(as.character(act$interval))
act$date <- as.Date(act$date)
```

### QUESTION 2| Histogram of the total number of steps taken each day.

I used the following packages. "In this case I hide the result/message to avoid unnesasary data": 

```{r warning=FALSE,message=FALSE}
library(dplyr)
library(lubridate)
library(ggplot2)
```

First, I getting the information per day by groupon by "date" and summurize per "steps".

```{r}
byday<- act %>% group_by(date) %>%
      summarize(steps=sum(steps))

```

Then, I plotted the graphic with 'date' variable on the X and 'steps' on the Y axis. 

```{r}
barplot(byday$steps,names.arg=byday$date, xlab = "Days", ylab = "Steps", main = "Number of steps taken each day",)
```

### QUESTION 3| Mean and median number of steps taken each day.

Using the 'byday' data frame created on question 2, we can calculate mean and median of steps per day:   

```{r}
mean(byday$steps,na.rm = TRUE)
median(byday$steps,na.rm = TRUE)
```
### QUESTION 4 | Time series plot of the average number of steps taken.

First, I getting the average steps per day (mean) by groupon by "date" and summurize per "steps": 
```{r}
meanday<- act %>% group_by(date=floor_date(date, "day")) %>%
      summarise(steps=mean(steps))
```
I used ggplot2 to plot a time series graphic showing the mean steps per day:  

```{r}
tseries <- ggplot(meanday, aes(x=date, y=steps))+
   geom_line(color="#69b3a2")+ 
   geom_point()+
   xlab("Days")+
   ylab("Steps")+
   ggtitle("Average number of steps taken each day") 
tseries
```

### QUESTION 5 | The 5-minute interval that, on average, contains the maximum number of steps.

Again, I used dplyr package to group steps by intervals and see the average number of steps of each one. Then I filter:  
```{r}
inter<- act %>% 
      filter(steps, interval) %>% 
      group_by(interval) %>%
      summarise(steps=mean(steps)) %>%
      arrange(desc(steps))
     
inter[1,]
```
### QUESTION 6 | Code to describe and show a strategy for imputing missing data.

First, launch "mice" Packages.
 
```{r warning=FALSE,message=FALSE} 
library(mice)
```
Then, I run the mice function to impute values for missing numbers and check the 3 different imputation options to see is the imputaiton that makes more sence. 

```{r} 
impute <- mice(act[,1:3], m=3, seed = 123)
```

After that, I replaced the NA for the imputation values and compare the new data set act2, with the origina act. 

```{r} 
act2<- complete(impute,1)
summary(act2)
summary(act) 
``` 

Finally, we see above that act2 does not contain NA values. Below I plot a graph showing the  imputation values in a different color.

```{r} 
xyplot(impute, steps~date | .imp, pch = 20, cex = 1.2) 
``` 


### QUESTION 7 | Histogram of the total number of steps taken each day after missing values are imputed.

I created a new vector "baday2" with no mission values "taken the new data frame act2". Then I grouped it by 'date' and sumaridez by 'steps'.    
```{r}
byday2 <- act2 %>% group_by(date) %>% 
           summarize(steps=sum(steps))
```
Create a new graph with no NA values:  

```{r}           
barplot(byday2$steps,names.arg=byday2$date, xlab = "Days", ylab = "Steps", main = "Number of steps taken each day",)          
```
### QUESTION 8 | Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends.

```{r}   
act2$date <- as.Date(act2$date)
act2$steps <- as.numeric(as.character(act2$steps))
act2$interval <- as.numeric(as.character(act2$interval))
```

First, I created a new variable called wDay with the two levels ('weekend', 'weekday').

```{r}   
weekdays1 <- c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday')
act2$wDay <- factor((weekdays(act2$date) %in% weekdays1), levels=c(FALSE, TRUE), labels=c('weekend', 'weekday')) 
```
Then, I group the data by 'interval' and summarise by 'steps'.

```{r}   
alldays<- act2 %>% 
      group_by(interval, wDay) %>%
      summarise(steps=mean(steps)) 
```
Finally, I plotted a 2 graphic to compare 'weekend' and 'weekday' intervals.
```{r} 
wdayplot <- ggplot(alldays, aes(x=interval, y=steps)) + 
   geom_line(color="#69b3a2") + 
   geom_point()+
   facet_grid(.~wDay)+ 
   xlab("Intervals")+
   ylab("Steps")+
   ggtitle("Average weekdays vs. weekedns steps per interval")  

wdayplot

```



            




